Sketch Recognition using Particle Swarm Optimizations

### 2009-05-30 paper2/Paper2.tex
The beginig of satureday changes i did on thursday in ERI

\author[ERI]{Maha El Meseery\footnote{Crossponding author}}
\author[ERI]{Mahmoud Fakhr El Din}
\author[ERI]{Samia Mashali}
\author[FAC]{Magda Fayek}
\author[FAC]{Nevin Darwish}
\address[ERI]{Signals Processing Group, Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
\address[FAC]{Computer Engineering Department\\
	Faculty of Engineering,Cairo University\\Cairo, Egypt\\
	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{abstract}
%%This study 
Sketches and drawings are widely used as simple methods of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to form symbols which are further identified using a Support Vector Machines (SVM) classifier. Two versions of PSO segmentation algorithms were implemented. Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols. Results show that both PSO algorithms produce better segmentation than most current systems. This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested. The results prove that the system is effective specially in datasets with low diversity in shapes (i.e. shapes are very simillar) as in Logic Design dataset (LD-DS). %The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS). %Result show that the system is especially effictive on dataset with % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.

@@

%\begin{abstract}
%%This study 
%Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).  % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to %formulate symbols which are further identified. The results show that using PSO improves segmentation results %which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.  
%\end{abstract}

\section{Introduction}

@@

Sketch recognition can be based on either Pattern Recognition (PR) method or Artifical Intelligance (AI) based method. In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them. Systems that employ such representation must first pre-process. strokes to break them into geometric primitives. Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes network, and constrain statisfaction. A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}.

Recognition based on AI methods is applied in \cite{SketchRead2007}. The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols. Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch. 
%The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.  
%Gestures are pen strokes drawn by the user that can immediately recognized by the system.
%role of which each stroke plays in the structure of the sketched symbol is used to determine shapes 
Systems that uses Pattern Recognition methods compute properties that represent different symbols then use a classification method to recognize symbols. The properties used can either be based on gesture properties, global shape properties or image based properties. Gestures properties are set of features that represent how the user draws a stroke not how it look like. In gestures properties approcah, the systems uses the role of each stroke in the structure of the symbol determine the shapes. System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12}. In global shape properties approach, systems investigate general shape properties and its underlying strokes. Using shape properties (i.e. ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}. Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes. Image based methods are used in \cite{Oltmans07} where they computed a code book of 1000 features to recognize symbols.

Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}. This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties. \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.

In this paper we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches. Our method tries to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system use. Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced. The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.

@@

\section{Discrete Particle Swarm Optimization}
\label{sec:ParticleSwarmAlgorithm}
%\section{Particle Swarm Algorithm}
%\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}. Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}
where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
\label{eq:Swarm}
\end{equation}
 where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
 After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}.  
\begin{equation}
   P(i)\Leftarrow 
\{
\begin{array}{c} 
1 \quad \quad if\quad r_{3}>p_{i}  \\

0 \quad \quad if\quad r_{3}<p_{i} 
\label{eq:descrite}
\end{array}\}
\end{equation}
 where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.

\section{The Proposed System}
\label{Sysdisc}
 The suggested sketch recognition system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.

\subsection{Preprocessing}
\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
 Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}. It was noted that time difference provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mention that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals as long as the pointing device is moving. Therefore, there are no new detected points when the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}). Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.  Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
 \begin{figure}[h]
	\centering
		\includegraphics[scale=0.5]{images/stroke3.jpg}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}
 \begin{figure}
	\centering
			\subfigure[ Speed Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
			\hfill
			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
	\caption{Speed and Time Difference Graphs}
	\label{fig:speed2Distance}
\end{figure}
Dominant points are characterized by lower speed values, higher curvature and direction values. However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information. Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.

After the system computes the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvature. However, using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information  is calculated. This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(R_i)$ of each region $R_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$ as the preprocessing. It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage. % (as shown are redundant)
\begin{figure}
	\centering
\includegraphics[scale=0.7]{images/ppd.jpg}
	\caption{Possible Dominante Points}
	%\label{fig:ppd}
	\label{fig:ppd999}	
\end{figure}

\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves. After two different segmentations are generated, the system chooses the segmentation with the minimum error. The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke. % The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing. The output is a set of dominant points which are connected by either lines or curves.

%///AllBlockDiagram.jpg  you can change the block to the main block. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.48]{images/blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting}
%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.  
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

\begin{equation}
E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
\label{eq:circleFit}
\end{equation}

 where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke. A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.  After few iteration, the final fit error of the estimated ellipse is reported. Another measure is used to compute the efficiency of the final estimated ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered. This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.

 \begin{equation}
eff= (P_{percent}/E)
\label{eq:circleError}
\end{equation}
 \begin{equation}
P_{percent}  = L_{stroke} /P_{ellipse} 
\label{eq:ErrorArea}
\end{equation}
where $E$ is the error computed by Equation(\ref{eq:circleFit}), $L_{stroke}$ total length of stroke and $P_{ellipse} $ is the perimeter of the estimated ellipse. If $eff$ is more than threshold $th_{Ellipse}$\footnote{By trial and error best threshold found was $th_{Ellipse}=0.2$} then the stroke is segmented as an ellipse otherwise the system proceeds to the next step.

\subsection{Non ellipse fitting algorithm}
\label{subsubsec:Discreteparticleswarmalgorithm}
\begin{figure}
	\centering
		%\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{images/ppd.jpg}}
	%	\hfill
	 \includegraphics[scale=0.7]{images/pso1.jpg}			
	\caption{DPSO algorithm encoding}% a) Possible dominate point b) Particle encoding  
	\label{fig:pso1}
\end{figure}
Two DPSO algorithms are used to generate different segmentations for each stroke the user draws. For each segmentation generated an error is evaluated. The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}). The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions. 
\begin{description}
	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.

>>>>

\author[ERI]{Maha El Meseery\footnote{Crossponding author}}
\author[ERI]{Mahmoud Fakhr El Din}
\author[ERI]{Samia Mashali}
\author[FAC]{Magda Fayek}
\author[FAC]{Nevin Darwish}
\address[ERI]{Signals Processing Group, Computers and Systems Department, 
  Electronic Research Institute, Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
\address[FAC]{Computer Engineering Department,
	Faculty of Engineering,Cairo University, Cairo, Egypt\\
	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{abstract}
%%This study 
Sketches and drawings are widely used as simple methods of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to form symbols which are further identified using a Support Vector Machines (SVM) classifier. Two versions of PSO segmentation algorithms were implemented. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that both PSO algorithms produce better segmentation than most current systems. This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested. The results prove that the system is effective specially in datasets with low diversity in shapes, i.e., shapes are very simillar, as in Logic Design dataset (LD-DS). %The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS). %Result show that the system is especially effictive on dataset with % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.

@@

%\begin{abstract}
%%This study 
%Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).  % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to %formulate symbols which are further identified. The results show that using PSO improves segmentation results %which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.  
%\end{abstract}
\pagenumbering{arabic}
\section{Introduction}

@@

Sketch recognition can be applied either using  Pattern Recognition (PR) based method or Artifical Intelligance (AI) based method. In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them. Systems that employ such representation must first pre-process. strokes to break them into geometric primitives. Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes belief network, and constrain statisfaction. A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,SketchRead2007}.  %Ladder30,AlvaradoFreedom42

Recognition based on AI methods as applied in \cite{SketchRead2007} has the principle problem of computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to be segmented and grouped to create reliable fragment of geometric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols. Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch. This is because they assume that each stroke plays a role in the structure of the shape. 
%The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.  
%Gestures are pen strokes drawn by the user that can immediately recognized by the system.
%role of which each stroke plays in the structure of the sketched symbol is used to determine shapes 
%. discribing the role  how user draws individal storkes  properties that discribe how users draw individual strokes in drawing the symbol.  to . The identification of individual stroke is used to determine  The symbols are recognized by identifing the individual strokes.
In PR based methods, systems compute properties that discribe symbols then use a classification method to recognize symbols. The properties used can either be based on gesture properties, global shape properties or image based properties. Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a regtangle when a user draws a L corner). In gesture properties approcah, systems identify symbols by using gesutre properties of some of its  underlying strokes. System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12,aideddesgin22}. In global shape properties approach, systems investigate general shape properties and its underlying strokes. Using shape properties (e.g., ratio of the area of convex hull to the area of shape) relaxes the assumption and the role of each stroke in the symbol but it does not truly represent the appearance or fine details of the symbol \cite{DiagramOfflineConvexHull,Cali63}. Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes. These system convertes the shapes into image and extracte image based features (e.g. local gradient feature)\cite{Oltmans07,imagetrainable48}. In \cite{imagetrainable48}, they used four different distance measures to to match user symbol with the prototype database. The prototype of a shape is represented as an image of the symbol that is normalized and scaled.  However, the problem is that the shape prototype must represent all transformation and variation the input shape can have.   %Image based methods are used in \cite{Oltmans07,imagetrainable48} where they computed a code book of 1000 features to recognize symbols.

Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}. This assumption allows the system to temporally group strokes and then recognizes them based on their strokes or global shape properties. Alvarado \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.

In this paper, we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches. Our method attempts to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system uses. Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced. The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.

@@

\section{Discrete Particle Swarm Optimization}
\label{sec:ParticleSwarmAlgorithm}
%\section{Particle Swarm Algorithm}
%\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
%The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution in the solution space with a particle\cite{PSOFirst}. Each particle moves the n with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.

\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}
%where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
\label{eq:Swarm}
\end{equation}
 where  $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.

 After each iteration the global best $g_{best}$ particle and the agent local best particle $l_{best}$are evaluated based on the maximum fitness functions of all particles in the solution space. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}). The \textit{DPSO} handles particle values of either $0$ or $1$ \cite{PSODisceret}.  
\begin{equation}
   P(i)\Leftarrow 
\left \{
\begin{array}{c} 
1 \quad \quad if\quad r_{3}>p_{i}  \\

0 \quad \quad if\quad r_{3}<p_{i}, 
\label{eq:descrite}
\end{array}\right.
\end{equation}
 where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.

\section{The Proposed System}
\label{Sysdisc}
 The suggested sketch recognition system is divided into three main steps; 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.

\subsection{Preprocessing}
\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
 Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}. Figure \ref{fig:orignalStroke} shows an example of an input stroke drawn by users.  Figure \ref{fig:speed2Distance} shows the graph of time difference and the speed for the stroke in Figure \ref{fig:orignalStroke}. It was noted that time difference provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mention that the pointing device (e.g., mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals as long as the pointing device is moving. Therefore, there are no new detected points when the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}). Contrary to time difference information, the speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data (Figure \ref{fig:speed2Distance}).  %Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}

 \begin{figure}[h]
	\centering
		\includegraphics[scale=0.5]{images/stroke3.jpg}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}

 \begin{figure}
	\centering
			\subfigure[Speed versus Displacement Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
			\hfill
			\subfigure[ Time Difference  versus Displacement Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
	\caption{Speed and Time Difference Graphs}  a) speed graph has lot of noise in the data with respect to b) time difference.   The black points denote the location of the possible dominant points $P_{pd}$. Black rectangles represent regions $R_i$ that in a) are lower than the thershold. b) are above the thershold. 
	\label{fig:speed2Distance}
\end{figure}
%is calculated as the angle between two vectors and  where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point $P_{i - 1}$ to point $P_i$ 
Dominant points are characterized by low speed values, high curvature and direction values. However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information. Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ calculated as the angle between vectors $\overrightarrow {P_{i - 1} P_i }$ and the $x-axis$ and the curvature is considered as the change in direction $d$ with respect to length $s$ i.e. $c= \Delta d/\Delta s$.

After the system computes the speed, time difference and curvature information, it proceeds to detect the points that have low velocity and high curvature. However, using simple differentiation to detect local extreme points results in false points due to the non smooth curves. Hence, our system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information  is calculated. This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(R_i)$ of each region $R_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$ in the preprocessing phase. It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage. % (as shown are redundant)
\begin{figure}
	\centering
\includegraphics[scale=0.7]{images/ppd.jpg}
	\caption{Possible Dominante Points} An illustration of stroke points and the expected \textit{possible dominant points} $P_{pd}$.  
	%\label{fig:ppd}
	\label{fig:ppd999}	
\end{figure}

\subsection{Segmentation}
\label{seg}
In the segmentation stage, the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}. The ellipse fitting step mainly helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves. Hence we generate two different segmentations, the system chooses the segmentation with the minimum error. The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke. % The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing. The output is a set of dominant points which are connected by either lines or curves. 
%After two different segmentations are generated, 
%///AllBlockDiagram.jpg  you can change the block to the main block. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.48]{images/blockSmall.jpg}
	\caption{Segmentation Block} 
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting}
%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov}.  
This process attempts to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ and $height$ are the width and height of the bonding box. The least square fitting algorithm \cite{chernov} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})  
\begin{equation}
E = \sum\limits_{i = 0}^N {\frac{{(x_i - x_0 )}}{{a^2 }}^2  + \frac{{(y_i - y_0 )}}{{b^2 }}^2  - 1} 
\label{eq:circleFit}
\end{equation}
where $N$ is number of points in the stroke, $a$ \& $b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point and $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke. A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.  After few iterations, the final fit error of the estimated ellipse is reported. Another measure is used to compute the efficiency of the final estimated ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered. This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or a full ellipse.
